\hydrogen{} and \carbon{} Network Strategies for the European Energy System

### 2023-12-29 report/submissions/journal/paper/paper.tex
integrate final internal review suggestions

\documentclass[twocolumn]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{cuted}  % Add the cuted package
\usepackage{caption}
\captionsetup{font=small}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{fancyhdr,lipsum}
\usepackage{subcaption}
\usepackage[shortcuts]{extdash}

@@

@@

At the same time, collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
% Some natural gas pipelines have already been repurposed to transport hydrogen~\cite{RohrFreiFuer}.
Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer to purchase \carbon{} on the other hand. The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \carbon{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential~\cite{ToolsGreenTransition} of 1.7 Gt carbon annually from point sources, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}. However, it's important to note that carbon sequestration technology, unlike carbon capture, is not yet fully mature and its full potential and implications remain to be explored in depth.
To advance the carbon economy, the Clean Air Task Force emphasizes the importance of developing a carbon transport system in Europe alongside a hydrogen network~\cite{lockwoodEuropeanStrategyCarbon}. Carbon pipelines, considered a mature technology, have seen widespread installations in the United States and Canada, primarily to supply \carbon{} to enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.

@@

In the network layouts, the slight shift in CU deployment results in partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a). This adjustment, however, leaves 80\% of the hydrogen topology intact. Simultaneously, the \carbon{} network expands routes from Central and Eastern Europe to the shores (1.2~bn€/a). The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios. The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities. In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remain mainly untouched (for detailed cost-contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).

@@

Fabian Hofmann and Christoph Tries are funded by the BreakthroughEnergy Initiative~\cite{BreakthroughEnergy2023}.

@@

\clearpage
\subsection{Inter-sectoral \carbon{} flows}

The following figure show Sankey Diagrams of the carbon flows between the sectors.

\begin{figure}[h!]
    \centering
    \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} assuming net \carbon{} neutrality.}
    \label{fig:sankey_diagramm}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.8\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} with a net \carbon{} removal target.}
    \label{fig:sankey_diagramm_nn}
\end{figure}

>>>>

\documentclass[twocolumn]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{cuted}  % Add the cuted package
\usepackage{caption}
\captionsetup{font=small}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{hyperref} % Add this line to enable hyperlinks
\usepackage{fancyhdr,lipsum}
\usepackage{subcaption}
\usepackage[shortcuts]{extdash}

@@

\hypersetup{
colorlinks=true,
    linkcolor=black, % Change the link color to blue
    urlcolor=blue, % Change the URL color to blue
    citecolor=black, % Change the citation color to blue
    filecolor=black % Change the file color to blue
}

@@

At the same time, collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
% Some natural gas pipelines have already been repurposed to transport hydrogen~\cite{RohrFreiFuer}.
Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer to purchase \carbon{} on the other hand. The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \carbon{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential~\cite{ToolsGreenTransition} of 1.7 Gt carbon annually from point sources, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}. However, it is important to note that carbon sequestration technology, unlike carbon capture, is not yet fully mature and its full potential and implications remain to be explored in depth.
To advance the carbon economy, the Clean Air Task Force emphasizes the importance of developing a carbon transport system in Europe alongside a hydrogen network~\cite{lockwoodEuropeanStrategyCarbon}. Carbon pipelines, considered a mature technology, have seen widespread installations in the United States and Canada, primarily to supply \carbon{} to enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.

@@

In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a). This adjustment, however, leaves 80\% of the hydrogen topology intact. Simultaneously, the \carbon{} network expands routes from Central and Eastern Europe to the shores (1.2~bn€/a). The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios. The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities. In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remain mainly untouched (for detailed cost contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).

@@

Fabian Hofmann and Christoph Tries are funded by the Breakthrough Energy Initiative (\href{https://breakthroughenergy.org/}{https://breakthroughenergy.org/}).

@@

% \clearpage
% \subsection{Inter-sectoral \carbon{} flows}

% The following figure show Sankey Diagrams of the carbon flows between the sectors.

% \begin{figure}[h!]
%     \centering
%     \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
%     \caption{Sankey diagram of the carbon flows in the \hybridscenario{} assuming net \carbon{} neutrality.}
%     \label{fig:sankey_diagramm}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=.8\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
%     \caption{Sankey diagram of the carbon flows in the \hybridscenario{} with a net \carbon{} removal target.}
%     \label{fig:sankey_diagramm_nn}
% \end{figure}
